Crossing Origins by Crossing Formats

### 2012-11-07 main.tex
Updated section Crossing origins

Server side verfication

@@

A polyglot can be used for malicious purposes if there is a difference 
between the expected execution environment and the actual environment 
it is executed in.
In the situation where

@@

For this attack to be successful certain requirements needs to be 
fulfilled. 
\begin{itemize}
	\item The original content must include parameters under the control of the 
				user.
	\item The user-supplied content must circumvent any validation or filtering.
	\item There must exist a way of coercing the resulting polyglot to be 
				interpreted as the malicious format.
	\item The malicious format must have capabilities that are restricted or 
				missing in the benign format.
\end{itemize}

@@

Much like a cross-site scripting attack, we consider a scenario where a website serves documents which are based upon user-supplied content.

The same-origin policy is intended to prevent a document from one origin from accessing the content of a document of another origin. At the same time it allows a script-tag in one document to include JavaScript of another origin and execute it in the context of that document. If an attacker is able to inject an HTML-document with JavaScript syntax in such a way that it became valid JavaScript, the attacker could include the document across origins via a script-tag and extract its content, thereby circumventing the same-origin policy.

Consider the php-program in listing \ref{lst:php-sample}. Apart from being vulnerable to traditional cross-site scripting, it is vulnerable

By requesting the site with parameter $guess=</body></html>;x=<html><body>$, the resulting document, listing \ref{lst:html-js}, is both malformed but valid HTML and valid JavaScript in browsers that support E4X. An attacker can now include the page as a script in the attackers page and extract the secret from variable x.

\begin{lstlisting}[label=lst:php-sample,caption=Vulnerable PHP-program,captionpos=b]
<?php
session_start();
$_SESSION["secret"] = rand();
?>
<html><body>
Your guess: <?=$_GET["guess"]?>!
The secret is: <?=$_SESSION["secret"]?>
</body></html>
\end{lstlisting}
\begin{lstlisting}[label=lst:html-js,caption=Vulnerable PHP-program,captionpos=b]
<html><body>
Your guess: </body></html>;x=<html><body>!
The secret is: 521352342
</body></html>
\end{lstlisting}
\if 0
$
\fi

This attack has been suggested as a way to bypass Content Security Policy. As E4X is only supported in the SpiderMonkey JavaScript engine, used in Firefox, this limits the exploitability of this scenario.
While this is an unlikely scenario using JavaScript as the injected format, section \ref{sec:vuln} will show how this can be achieved in practice.

@@

% Description of attack technique
A content repurposing attack attacks a service that accepts and serves user supplied files. The files are verified to be of a certain benign file type, e.g. an image, before being served to the end user. By submitting a polyglot to such a service, an attacker can trick the service into serving malicious content under the origin of the service. The service can verify the

% Building blocks
For this attack to be successful certain requirements needs to be 
fulfilled.

\begin{itemize}
	\item The service must accept and serve user-supplied files.
	\item The service must verify the user-supplied polyglot as the benign 
				format.
	\item There must exist a way of coercing the polyglot to be 
				interpreted as the malicious format.
\end{itemize}

@@

@@

Why PDF format?

@@

The object-tag allows the developer to specify which format 
the embedded should be interpreted as.

@@

@@

@@

In the case of PDF; always encode tokens related to the PDF 
syntax. Generally, hard to ensure that tokens for all potential 
file formats are properly encoded without breaking anything.

>>>>

Server side verification

@@

A polyglot can be used for malicious purposes if there is a difference 
between the expected execution environment and the actual environment 
it is executed in. If the content is verified to be benign in 
In the situation where

@@

The target in this scenario is a user of the vulnerable website. The polyglot 
is served in the benign format on the vulnerable server. However, the attack 
is initiated from a page belonging to the attacker. While the user is logged 
in to the service that the vulnerable website provides, the user 
unsuspectingly visits the attackers malicious website. The page embeds the 
polyglot content, but coerces it to be interpreted as the malicious format. 
The attacker can now exploit the vulnerability, either by making requests to 
the vulnerable website in the context of the user; or by extracting 
information about the user that is stored on the vulnerable website.

In situations like these, the same-origin policy is intended to prevent a document from one origin from accessing the content of a document of another origin. However, depending on the capabilities of the malicious format used in the attack, it may be possible to circumvent the same-origin policy.

For this attack to be successful certain requirements needs to be 
fulfilled. 
\begin{itemize}
	\item The original content must embed external information under the control 
				of the attacker.
	\item The user-supplied content must circumvent any validation or filtering.
	\item There must exist a way of coercing the resulting polyglot to be 
				interpreted as the malicious format.
	\item The malicious format must have capabilities that are restricted or 
				missing in the benign format.
\end{itemize}

@@

%Much like a cross-site scripting attack, we consider a scenario where a 
%website serves documents which are based upon user-supplied information.

%The same-origin policy is intended to prevent a document from one origin from 
%accessing the content of a document of another origin. 
As an example the same-origin policy allows a script-tag in one document to 
include JavaScript of another origin and execute it in the context of that 
document. If an attacker is able to inject an HTML document with JavaScript 
syntax in such a way that it became valid JavaScript, the attacker could 
include the document across origins via a script-tag and extract its content, 
thereby circumventing the same-origin policy.

%Consider the php-program in listing \ref{lst:php-sample}. Apart from being 
%vulnerable to traditional cross-site scripting, it is vulnerable

%By requesting the site with parameter $guess=</body></html>;x=<html><body>$, 
%the resulting document, 
The document in listing \ref{lst:html-js}, is both malformed but valid HTML 
and valid JavaScript in browsers that support E4X. The document can be 
included as a script in another document and the secret can be extracted from 
variable x. This implies that if an attacker is able to inject HTML syntax 
into a document, such that the result is two balanced E4X trees, then 
information can be extracted across origins.

\begin{lstlisting}[label=lst:html-js, caption=HTML / E4X Polyglot]
<html><body>
Hello </body></html>;x=<html><body>!
Your secret is 'safe with me'.
</body></html>
\end{lstlisting}

This attack has been suggested as a potential way to bypass Content Security 
Policy \ref{required}. 
As E4X is only supported in the SpiderMonkey JavaScript engine, used in 
Firefox, this limits the exploitability of this example.
While this is an unlikely scenario using JavaScript as the injected format, 
section \ref{sec:vuln} will show how this can be achieved in practice.

% The below example is too long and complicated.

%\begin{lstlisting}[label=lst:php-sample,caption=Vulnerable PHP-program,captionpos=b]
%<?php
%session_start();
%$_SESSION["secret"] = rand();
%?>
%<html><body>
%Your guess: <?=$_GET["guess"]?>!
%The secret is: <?=$_SESSION["secret"]?>
%</body></html>
%\end{lstlisting}
%\begin{lstlisting}[label=lst:html-js,caption=Vulnerable %PHP-program,captionpos=b]
%<html><body>
%Your guess: </body></html>;x=<html><body>!
%The secret is: 521352342
%</body></html>
%\end{lstlisting}
%\if 0
%$
%\fi

@@

% Description of attack technique
A content repurposing attack attacks a website that allows and serves user supplied files. This could be a file sharing service, an image database, or a job search service. The files are verified to be of a certain benign file type, e.g. an image, before being served to the end user. By submitting a polyglot to such a service, an attacker can trick the service into serving malicious content under the origin of the service. The service will verify the file, assuming that it is of the benign format, but once the content is served an attacker can coerce the content to be interpreted as the malicious format.

% Building blocks

The difference in this scenario from the previous one is in the nature of the vulnerability. Where in the previous scenario an existing document was injected to become a polyglot, this scenario targets services which allow an attacker to introduce new documents.
As in the previous scenario, the target is a user of the vulnerable website.

For this attack to be successful certain requirements needs to be 
fulfilled.

\begin{itemize}
	\item The service must accept and serve user-supplied files.
	\item The service must verify the user-supplied polyglot under the 
				assumption that it is in the benign format.
	\item There must exist a way of coercing the polyglot to be 
				interpreted as the malicious format.
\end{itemize}

@@

In this section we give concrete examples of the attacks described in the previous section, using the PDF format as the running example.

@@

\subsubsection{The PDF format}

% Why PDF format?

The PDF format is a widely used, text based format with powerful capabilities. It can be used for displaying text, rendering graphics, animation and dynamic content. It is also a container format which allows for embedding of other files and formats. This and the fact that it is designed to be robust against errors, provides us with a format that is powerful, flexible and ideal for creating malicious polyglots.

%Several design choices

Furthermore the PDF format is of particular interest to us because it can execute JavaScript and send HTTP requests to the server from which the document originated.

@@

%The object-tag allows the developer to specify which format 
%the embedded content should be interpreted as.

The object-tag is used to embed content in a web page, which requires a plug-in to run. The plug-in is selected based on the MIME-type of the content, but because the server delivering the content might not be able to determine its format, the object tag allows a developer to set the type-attribute to guide the browser in which plug-in to run. When the type attribute is used, the corresponding plug-in is run regardless of which MIME-type the server returns and it is up to the plug-in to respond to and handle the content it is served.

In the case of the Adobe Reader plug-in completely disregards the MIME-type supplied by the server and will attempt to interpret any content as PDF.

@@

\subsection{Polyglot upload}

@@

This section will describe our recommendations for mitigation in each of the affected components.

@@

%In the case of PDF; always encode tokens related to the PDF 
%syntax. Generally, hard to ensure that tokens for all potential 
%file formats are properly encoded without breaking anything.

Preventing syntax injection attacks on the server-side poses severe challenges. Even server-side filtering of HTML syntax to prevent cross-side scripting attacks has proved difficult due to the many contexts in which JavaScript can be introduced. Filtering all potentially harmful tokens from all formats in which a document could be interpreted, will likely be near impossible.
